### 2020-01-04 paper/cover_letter.tex
Adding cover letter

\documentclass{letter}
\usepackage[left=1in,top=0.15in,twoside=false]{geometry}
\usepackage{color,hyperref}
\definecolor{darkblue}{rgb}{0.0,0.0,0.7}
\hypersetup{colorlinks,breaklinks,
            linkcolor=darkblue,urlcolor=darkblue,
            anchorcolor=darkblue,citecolor=darkblue}

% \signature{Agrim Gupta}
\address{From \\
Agrim Gupta\\
Department of Electrical  Engineering\\
Indian Institute of Technology Bombay, Mumbai\\
India\\
E-mail: agrim9gupta@gmail.com
\date{\today}
}

\begin{document}
\begin{letter}{To\\
Prof. M. Di Renzo\\
Editor-in-Chief\\
IEEE Communication Letters}
\vspace{0.5in}

\opening{Dear Prof. M. Di Renzo,}

Subject: \textbf{Submission of manuscript title ``Reservoir Computing Framework for Predictive Quantization of MIMO-OFDM SVD Precoders'' for potential publication in
  IEEE Communication Letters.}

Please find enclosed our manuscript titled ``Reservoir Computing Framework for Predictive Quantization of MIMO-OFDM SVD Precoders'' for review and potential
  publication in IEEE Communication Letters. The co-authors of this manuscripts are
\begin{enumerate}
\item  Pranav Sankhe

Department of Electrical  Engineering

Indian Institute of Technology Bombay, Mumbai

India

E-mail: pranavsankhe40@gmail.com

\item Prof. Kumar Appaiah

Department of Electrical  Engineering

Indian Institute of Technology Bombay, Mumbai

India

E-mail: akumar@ee.iitb.ac.in

\item Prof. Manoj Gopalkrishnan

Department of Electrical  Engineering

Indian Institute of Technology Bombay, Mumbai

India

E-mail: manojg@ee.iitb.ac.in

\end{enumerate}

Should you require additional details, please do e-mail me.

Thank you.
\vspace{0.3in}

Sincerely,

Agrim Gupta

\end{letter}
\end{document}